---
title: "Homework 4"
author: "Mari Sanders"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1 

a) 
Sign Test
```{r}
blood_sugar <- 120 - c(125,123,117, 123, 115, 112, 128, 118, 124, 111, 116, 
                       109, 125, 120, 113, 123, 112, 118, 121, 118, 122, 115,105, 
                       118, 131)

binom.test(sum(blood_sugar > 0), length(blood_sugar), alternative = "less")

```

The p-value is 0.7878, which means we would fail to reject the null hypothesis. 

b)
Wilcox Signed Rank Test
```{r}
blood_sugar2 <- c(125,123,117, 123, 115, 112, 128, 118, 124, 111, 116, 
                       109, 125, 120, 113, 123, 112, 118, 121, 118, 122, 115,105, 
                       118, 131)
wilcox.test(blood_sugar2, mu = 120, alternative = "less")
```

# Problem 2 

a) 

```{r}
brain <- read_xlsx("data/Brain.xlsx") %>% janitor::clean_names() 
brain %>% 
  slice(-1) %>% 
  ggplot(aes(x = ln_brain_mass, y = glia_neuron_ratio)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  geom_point(color = "red") +
  geom_point(aes(x = brain$ln_brain_mass[1], 
                 y = brain$glia_neuron_ratio[1])) +
  guides(color = "none") +
  theme_classic()
human_ln_brain_mass <- brain$ln_brain_mass[1]

non_human <- lm(glia_neuron_ratio ~ ln_brain_mass, data = brain %>% slice(-1))

broom::tidy(non_human)

```

`glia_neuron_ratio` = 0.101 + 0.197*`ln_brain_mass`

b) 

```{r}
predict(non_human, newdata = data.frame(ln_brain_mass = human_ln_brain_mass))
```

This means that human `glia_neuron_ratio` is 1.471458. 


c) Since we want to look at humans compared to nonhuman primates, the interval for predicted mean glia-neuron ratio is more relevant. 

d) 

```{r}

predict(non_human, newdata = data.frame(ln_brain_mass = human_ln_brain_mass), interval = "confidence", level = 0.95)
```

We are 95% confident that the predicted mean glia-neuron ratio is between 1.229558 and 1.713358. 
e) 

The position of the human data point might pull the regression line to it and might not accurately predict the rest of the data points. 

# Problem 3 

a) 

```{r}
heart_disease <- read_csv("data/HeartDisease.csv") 
```

This data has `nrow(heart_disease)` rows and `ncol(heart_disease)` columns. It includes `names(heart_disease)` variables. The main outcome is `ERvisits` and the main predictor is `total cost (in dollars)`. The covariates might be `age`, `complications`, `gender`, and `duration`.

```{r}
heart_disease %>% 
  summarize(
    mean_totalcost = mean(totalcost, na.rm = TRUE),
    median_totalcost = median(totalcost, na.rm = TRUE),
    sd_totalcost = sd(totalcost, na.rm = TRUE),
    mean_ervisits = mean(ERvisits, na.rm = TRUE),
    sd_ervisits = sd(ERvisits, na.rm = TRUE),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    mean_complications = mean(complications, na.rm = TRUE),
    sd_complications = sd(complications, na.rm = TRUE),
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>% knitr::kable()
```

b) 

```{r}
heart_disease %>% 
  ggplot(aes(x = totalcost)) + geom_histogram()

heart_disease %>% 
  ggplot(aes(x = log(totalcost))) + geom_histogram()

```

Doing a log transformation seems to make the data look more normal. 

c) 

```{r}
heart_data <- 
  heart_disease %>% 
  drop_na() %>% 
  mutate(comp_bin = ifelse(complications > 0, 1, 0), 
         log_totalcost = log(totalcost)) %>%
  filter(!is.na(totalcost) & !is.infinite(totalcost) & 
         !is.na(log_totalcost) & !is.infinite(log_totalcost))
```

d) 

```{r}
heart_data %>% 
  ggplot(aes(x = log_totalcost, y = ERvisits)) + geom_point() + geom_smooth(method = lm, se = FALSE)
model <- lm(ERvisits~log_totalcost, data = heart_data)
broom::tidy(model)
```

`ERvisits` changes by 0.452 as `log_totalcost` increases by one unit. 

`log_totalcost` is signfiicant but the intercept, which is `ERvisits` is not significant because the p-value is greater than 0.05. 

e) 

```{r}
mlr <- lm(log_totalcost ~ ERvisits * comp_bin, data = heart_data)

broom::tidy(mlr)
```
`comp_bin` is significant, as well as `ERvisits` and the intercept `log_totalcost`, but the interaction between `ERvisits` and `comp_bin` is not significant. 

Since `comp_bin` is not significant, it is not an effect modifier. 

```{r}
mlr_nocomp <- lm(log_totalcost ~ ERvisits, data = heart_data)
broom::tidy(mlr_nocomp)
```

